---
title: "TEST Country Insights"
output: html_document
date: "2024-12-06"
toc: true
runtime: shiny
---

Testing a way to generate country-specific info and figures, creating ideally a downloadable report, this might work better as an html instead, if we wanted any kind of interactvity 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r}
country <- "Country" #specifiy country 
territory <- "Territory" #specify territory #NOTE -- how do we treat territory when a country's report is downloaded? maybe just a list of them 


#read in all data 
carbon_stocks <- read_csv("/carbon_stocks.csv")
emissions_factors <- read_csv("/emissions_factors.csv")
ecosystem_area <- read_csv("/ecosystem_area.csv")
citations <- read_csv("/citations.csv")


```

One idea of how to do this -> create lookup tables of each of the values we want to pull, by country, then use chunks of code to write text and include inline code to change these 
^ the one problem here is how do we change the country 

```{r}
#get top line numbers for country 
#made up numbers to test 
lowC <- 10
highC <-99
CI <-5
totalStock <- 150
mangroveClow <- 15
mangroveChigh <- 86


```


```{r} 
# Citations 
library(knitr)
 
#pull all country-level citations from big table 
citations
country_citations <- citations %>% filter(country == country)

knitr::kable(country_citations, caption = "Citations")

```



## Country Insights 

This report includes country-specific insights and more detailed analysis and figures, including carbon stocks, emissions factors, and ecosystem wetland area for mangrove, marsh, and seagrass ecosystems. This report details information for `r country`, including data from `r territory`



##MAP - Add interactive leaflet widget of country, pull from use in app 
```{r}
library(leaflet)

```




## Carbon Stock Estimates 

We estimate that `r country` and `r territory` contains between (`r lowC` to `r highC`, `r CI`) metric tonnes of soil C to a depth of 1 m, with a mean estimate of (total stock) metric tonnes C. 

This total estimate includes mangrove carbon stocks (low to high, 95% CI, mean = total stock),
tidal marsh carbon stocks

Seagrass carbon stocks were not included in the total value due to lack of a global, transparent, and independently assessed seagrass habitat map, however, best available areas and stocks are listed below.


## Wetland Areas and Activities Data 



## Including Plots

Include all plots from tabs in app 

```{r pressure, echo=FALSE}
plot(pressure)
```





