---
title: "US Blue Carbon Data Inventory"
author: "Coastal Carbon Network"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    theme: spacelab
    highlight: tango
---

```{r setup, include=FALSE}
# this sets the working directory to start where the R project is located
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

# no warnings or messages
# knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)

```

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Assemble all metrics in one table
v1 <- v1_metrics %>% 
  select(-c(contains("_weight"), contains("_rank"), contains("mean_"))) %>% 
  rename(spatial_metric = spatial_representativeness_metric) %>% 
  mutate(habitat_metric = ifelse(state == "PA" | state == "DC", NA, habitat_metric))
colnames(v1)[2:length(names(v1))] <- paste0(colnames(v1)[2:length(names(v1))], "_2022")

v2 <- v2_metrics %>% 
  select(-c(contains("_weight"), contains("_rank"), contains("mean_"))) %>% 
  rename(spatial_metric = spatial_representativeness_metric) %>% 
    mutate(habitat_metric = ifelse(state == "PA" | state == "DC", NA, habitat_metric))
# add v2 to distinquish these metrics in the wide format
colnames(v2)[2:length(names(v2))] <- paste0(colnames(v2)[2:length(names(v2))], "_2023")

all_metrics <- full_join(v1, v2) %>% 
  mutate(across(everything(), replace_na, 0)) %>% 
  mutate(quant_diff = quantity_metric_2023 - quantity_metric_2022,
         qual_diff = quality_metric_2023 - quality_metric_2022,
         spatial_diff = spatial_metric_2023 - spatial_metric_2022, 
         hab_diff = habitat_metric_2023 - habitat_metric_2022)

```

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Prepare spatial data for maps of composite scores
# Common projection for states
states_aea <- st_transform(states, crs = "+proj=aea +ellps=WGS84 +lat_1=29.5 +lat_2=45.5 +lon_0=-96 +x_0=0 +y_0=0")

# filter for coastal states
coastal_states <- states_aea %>% 
  # filter(STUSPS %in% unique(all_metrics$state)) %>% 
  rename(state = STUSPS) %>% 
  full_join(all_metrics) %>% 
  full_join(scores_v1 %>% filter(metric == "Total") %>% select(-metric) %>% rename(normalized_rank_v1 = normalized_rank)) %>% 
  full_join(scores_v2 %>% filter(metric == "Total") %>% select(-metric) %>% rename(normalized_rank_v2 = normalized_rank)) 

## Map States with Composite Scores ####

# generate bounding box
b <- st_bbox(coastal_states %>% filter(state %in% unique(all_metrics$state)))

```

## Methods

The methods for calculating data inventory metrics and generating composite scores are identical to those used in the first [Blue Carbon Data Inventory Report](https://smithsonian.github.io/CCRCN-Pew-Project/methods.html). The figure from the original analysis has been revised to reflect the updates that were applied to existing datasets. See documentation from [CCN Database Version 1.1.0](https://github.com/Smithsonian/CCN-Data-Library/releases) for full list of updates. All analyses were run using the most current version of the CCN database (V1.1.0). In order to recalculate the metrics from the initial inventory, this dataset was filtered to only include the data which were present in the [previous database version](https://doi.org/10.25573/serc.21565671.v1) (V1.0.0).

***

## Data Inventory 2022

### Report Card

```{r report-2022, fig.height=12, fig.width=12}

state_rank <- scores_v1 %>%
  filter(metric == "Total") %>%
  mutate(normalized_rank = ifelse(is.na(normalized_rank), 5, normalized_rank)) %>%
  arrange(-normalized_rank)

scores_v1 %>%
  mutate(metric = factor(metric, c("Total", "Quantity", "Quality", "Spatial coverage", "Habitat coverage"))) %>%
  mutate(state = factor(state, state_rank$state)) %>%
  mutate(normalized_rank = as.character(normalized_rank)) %>%
  ggplot(aes(metric, state, fill = normalized_rank)) +
  geom_tile(color = "white", size = 0.5) +
  # coord_equal() +
  # geom_hline(aes(yintercept = 6.5), size = 0.5, col = "grey") +
  # geom_hline(aes(yintercept = 9.5), size = 0.5, col = "grey") +
  # geom_hline(aes(yintercept = 17.5), size = 0.5, col = "grey") +
  # geom_hline(aes(yintercept = 21.5), size = 0.5, col = "grey") +
  geom_vline(aes(xintercept = 1.5), size = 0.5, col = "black") +
  ylab(NULL) + xlab(NULL) +
  scale_fill_brewer(palette = "RdYlBu", direction = -1,
                    labels = c("Best", "", "", "Fair", "", "", "Poor", "No Data")) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) + # wrap text

  # facet_wrap(vars(metric)) +
  theme_bw(base_size = 15) +
  theme(legend.title = element_blank(),
    # axis.text.x = element_text(angle = 90),
        axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) +
   ggtitle("State-Level Blue Carbon Data Report Card")

ggsave("blue_carbon_data_inventory/figures/report_card_2022.jpg", width = 6, height = 6)

```

### Composite Score Map

```{r fig.width=12, fig.height=6}
# ggplot score map V1
ggplot() + 
  # states
  geom_sf(data = coastal_states, mapping = aes(fill = as.character(normalized_rank_v1)), 
          color = "darkgrey",
          size = 0.1) +
  # custom colors...this scales it to 5 categories instead of 7, mismatching the colors
  scale_fill_manual(values = c("#4375B5", "#91BEDB", "#E0F3F8", "#FFFDC0", "#FEE090"), na.value = "lightgrey",
                    name = "Composite Rank", labels = c("Best", "", "Fair", "", "Poor", "No Data")) +
  # project aes w bounding box
  coord_sf(xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"]),
           crs="+proj=aea +ellps=WGS84 +lat_1=29.5 +lat_2=45.5 +lon_0=-96 +x_0=0 +y_0=0") +
  # Probably want to take off x and y axes text
  # theme_minimal() +
  theme_minimal(base_size = 15) +
  # ggtitle("Coastal State Composite Scores") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggsave("blue_carbon_data_inventory/figures/composite_score_map_2022.jpg", width = 7, height = 5)


# variation which has a more extreme red to blue scale, but this doesn't match the report card total score colors
# # ggplot score map V2
# coastal_states %>%
#   ggplot(aes(fill = as.character(normalized_rank_v2))) +
#   geom_sf() +
#   # scale_fill_continuous(type = "viridis") +
#   # scale_color_continuous() +
#   scale_fill_brewer(palette = "RdYlBu", direction = -1, na.value = "lightgrey",
#                     name = "Composite Rank", labels = c("Best", "", "Fair", "", "Poor", "No Data")) +
#   # project aes w bounding box
#   coord_sf(xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"]),
#            crs="+proj=aea +ellps=WGS84 +lat_1=29.5 +lat_2=45.5 +lon_0=-96 +x_0=0 +y_0=0") +
#   # Probably want to take off x and y axes text
#   theme_minimal() +
#   # ggtitle("Change in the Quantity of Available Blue Carbon Data") +
#   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

## Data Inventory 2023

### Report Card

```{r report-2023, fig.height=12, fig.width=12}

state_rank2 <- scores_v2 %>%
  filter(metric == "Total") %>%
  mutate(normalized_rank = ifelse(is.na(normalized_rank), 5, normalized_rank)) %>%
  arrange(-normalized_rank)

scores_v2 %>%
  mutate(metric = factor(metric, c("Total", "Quantity", "Quality", "Spatial coverage", "Habitat coverage"))) %>%
  mutate(state = factor(state, state_rank2$state)) %>%
  mutate(normalized_rank = as.character(normalized_rank)) %>%
  ggplot(aes(metric, state, fill = normalized_rank)) +
  geom_tile(color = "white", size = 0.5) +
  # coord_equal() +
  # geom_hline(aes(yintercept = 6.5), size = 0.5, col = "grey") +
  # geom_hline(aes(yintercept = 9.5), size = 0.5, col = "grey") +
  # geom_hline(aes(yintercept = 17.5), size = 0.5, col = "grey") +
  # geom_hline(aes(yintercept = 21.5), size = 0.5, col = "grey") +
  geom_vline(aes(xintercept = 1.5), size = 0.5, col = "black") +
  ylab(NULL) + xlab(NULL) +
  scale_fill_brewer(palette = "RdYlBu", direction = -1,
                    labels = c("Best", "", "", "Fair", "", "", "Poor", "No Data")) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) + # wrap text
  # facet_wrap(vars(metric)) +
  theme_bw(base_size = 15) +
  theme(legend.title = element_blank(),
        # axis.text.x = element_text(angle = 90),
        axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) +
   ggtitle("State-Level Blue Carbon Data Report Card")

ggsave("blue_carbon_data_inventory/figures/report_card_2023.jpg", width = 6, height = 6)
```

### Composite Scores Map

```{r fig.width=12, fig.height=6}

# ggplot score map
ggplot() + 
  # states
  geom_sf(data = coastal_states, mapping = aes(fill = as.character(normalized_rank_v2)), 
          color = "darkgrey",
          size = 0.1) +
  # custom colors...this scales it to 5 categories instead of 7, mismatching the colors
  scale_fill_manual(values = c("#4375B5", "#91BEDB", "#E0F3F8", "#FFFDC0", "#FEE090"), na.value = "lightgrey",
                    name = "Composite Rank",
                    labels = c("Best", "", "Fair", "", "Poor", "No Data")) +
  # project aes w bounding box
  coord_sf(xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"]),
           crs="+proj=aea +ellps=WGS84 +lat_1=29.5 +lat_2=45.5 +lon_0=-96 +x_0=0 +y_0=0") +
  # Probably want to take off x and y axes text
  theme_minimal(base_size = 15) +
  # ggtitle("Coastal State Composite Scores") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
        # legend.position = "none")

ggsave("blue_carbon_data_inventory/figures/composite_score_map_2023.jpg", width = 7, height = 5)

```

### Metric Score Overview

```{r fig.width=12, fig.height=12}

library(gridExtra)

## Construct State-Level Quadrat Figure ####
quantity_plot <- ggplot(data = all_metrics, aes(x = reorder(state, quantity_metric_2023), y = quantity_metric_2023)) +
  geom_point(pch=16) +
  geom_segment(aes(xend=state, yend=0)) +
  xlab(NULL) +
  ylab("Cores per 1000 ha tidal wetland") +
  # ggtitle("A. Quantity") +
  coord_flip() +
  # labs(color = "Detailed in report?") +
  theme_minimal(base_size = 15)

# (quantity_plot) 
ggsave("blue_carbon_data_inventory/figures/state_quantity_2023.jpg", width = 6, height = 6, quantity_plot)

quality_plot <- ggplot(data = all_metrics, aes(x = reorder(state, quality_metric_2023), y = quality_metric_2023)) +
  geom_point(pch=16) +
  geom_segment(aes(xend=state, yend=0)) +
  xlab(NULL) +
  ylab("High Quality Cores per 1000 ha \ntidal wetland") +
  # ggtitle("B. Quality") +
  coord_flip() +
  # labs(color = "Detailed in report?") +
  theme_minimal(base_size = 15) +
  theme(legend.position = "none")

# (quality_plot) 
ggsave("blue_carbon_data_inventory/figures/state_quality_2023.jpg", width = 6, height = 6, quality_plot)

spatial_plot <- ggplot(data = all_metrics, aes(x = reorder(state, spatial_metric_2023), y = spatial_metric_2023)) +
  geom_point(pch=16) +
  geom_segment(aes(xend=state, yend=0)) +
  xlab(NULL) +
  ylab("Mean Sampling Distance\n/ Idealized Sampling Distance") +
  # ggtitle("C. Spatial Representativeness") +
  coord_flip() +
  # labs(color = "Detailed in report?") +
  theme_minimal(base_size = 15) +
  theme(legend.position = "none")

# (spatial_plot) 
ggsave("blue_carbon_data_inventory/figures/state_spatial_2023.jpg", width = 6, height = 6, spatial_plot)


habitat_plot <- ggplot(data = all_metrics, 
                       aes(x = reorder(state, habitat_metric_2023), y = habitat_metric_2023)) +
  geom_point(pch=16) +
  geom_segment(aes(xend=state, yend=0)) +
  xlab(NULL) +
  ylab("1 - Distance Between Sampled\nand Mapped Proportions") +
  # ggtitle("D. Habitat Representativeness") +
  coord_flip() +
  # labs(color = "Detailed in report?") +
  theme_minimal(base_size = 15) +
  theme(legend.position = "none")

# (habitat_plot)
ggsave("blue_carbon_data_inventory/figures/state_habitat_2023.jpg", width = 6, height = 6, habitat_plot)

# assemble quadrat figure
grid.arrange(nrow=2, ncol=2,
             quantity_plot, quality_plot, 
             spatial_plot, habitat_plot,
             left = "Lower to Higher Ranking")

# # create object to save to figs folder
# grobs <- arrangeGrob(nrow=2, ncol=2,
#                      quantity_plot, quality_plot, 
#                      spatial_plot, habitat_plot,
#                      left = "Lower to Higher Ranking")
# 
# ggsave("blue_carbon_data_inventory/figures/metric_scores_2023.jpg", width = 6, height = 6, grobs)

```

### State Data Quality

This figure provides a more detailed look at the utility and completeness of the soil carbon data which is available for each state. The use category *Stocks* refers to soil cores which contain information that can be used to calculate carbon stock. For stocks, *good* indicates that carbon stock data is complete but not confirmed to represent the full wetland profile, and *best* indicates that carbon stock data is complete and represents a full profile. The *carbon accumulation rate (CAR)* category refers to soil cores wtih dated stratigraphy. In this category, *good* indicates the dating information is present but not complete, and *best* indicates that dating information is both present and complete. Lastly, the *Modeling* category refers to soil cores with both dated stratigraphy and associated elevation data, which is useful for modeling and forecasting wetland dynamics. In this category, *good* indicates that either (1) elevation data is low quality and dating information is present or (2) elevation data is high quality but dating information is incomplete. Soil cores that qualify for *best* in this category have high quality elevation data and complete dating information. See [Section 3.1](https://smithsonian.github.io/CCRCN-Pew-Project/methods.html#the-coastal-carbon-library) of the original data inventory report for more information.

Notes: These categories are not mutually exclusive, because a single profile may possess information which is relevant to all three use categories. Also note that Y-axes are scaled differently for each state. 

```{r quality-metric, echo=FALSE, fig.width=12, fig.height=12}

# Isolate quality codes from current core table and conver to long form
quality_smry <- v2_cores %>% 
  select(study_id, site_id, core_id, state, stocks_qual_code, dates_qual_code, elevation_qual_code) %>% 
  pivot_longer(cols = c(stocks_qual_code, dates_qual_code, elevation_qual_code), names_to = "use_category",
               values_to = "quality_code") %>% 
  drop_na(quality_code) %>% 
  distinct() %>% 
  count(state, use_category, quality_code, name = "core_count") %>% 
  # recode some values for plotting purposes
  mutate(use_category = recode(use_category,
                               "dates_qual_code" = "CAR",
                               "stocks_qual_code" = "Stocks",
                               "elevation_qual_code" = "Modeling"),
         quality_code = case_when(grepl("1", quality_code) ~ "Best",
                                  # quality_code == "A2" ~ "Better", # going to let this group in with "Good"
                                  grepl("2", quality_code) ~ "Good",
                                  grepl("3", quality_code) ~ "Good")) %>% 
  drop_na(state) %>% arrange(state, use_category)

# Plot the state breakdown

# order the tiers manually
# level_order <- c('Good', 'Better', 'Best')
level_order <- c("Stocks", "CAR", "Modeling")

quality_smry %>% 
  ggplot(aes(x = factor(use_category, level = level_order), core_count, fill = quality_code)) +
  geom_col() + 
  scale_fill_manual(values = c("#4375B5", "#91BEDB"), name = "Data Utility") +
  facet_wrap(~state, scales = "free_y", ncol = 4) +
  xlab("Use Category") + ylab("Number of Sediment Profiles") +
  theme_bw(base_size = 15) 

# values = c("#4375B5", "#91BEDB", "#E0F3F8", "#FFFDC0", "#FEE090")

# save the figure (needed some visual adjusting)
ggsave("blue_carbon_data_inventory/figures/state_data_utility_2023.jpg", width = 8, height = 6,
       quality_smry %>% 
         ggplot(aes(x = factor(use_category, level = level_order), core_count, fill = quality_code)) +
         geom_col() + 
         scale_fill_manual(values = c("#4375B5", "#91BEDB"), name = "Data Utility") +
         facet_wrap(~state, scales = "free_y", ncol = 4) +
         xlab("Use Category") + ylab("Number of Sediment Profiles") +
         theme_bw()
)

```

### State Habitat Representativeness

```{r fig.width=12, fig.height=12}

habitat_detailed <- habitatProportions(v2_cores) %>% 
  gather(key = "sampled_v_area", value = "proportion", -c(state, habitat_simplified)) %>% 
  mutate(sampled_v_area = recode(sampled_v_area, 
                                 "state_core_proportion" = "sampled",
                                 "mapped_proportion" = "estimated area")
         # habitat_simplified = recode(habitat_simplified, "seagrass, kelp and algal mats" = "estuarine aquatic bed")
         )

# create a custom color scale that will be used in the state reports as well
habs <- sort(unique(habitat_detailed$habitat_simplified))
hab_colorbrew <- c("#FC8D62", "#8DA0CB","#A6D854", "#FFD92F", "#E78AC3") # set 2 (colorblind friendly)
names(hab_colorbrew) <- habs # associate habitats with the colors
hab_colors <- scale_fill_manual(name = "habitat", values = hab_colorbrew) # create custom fill
# alternatives:
# hab_colorbrew <- brewer.pal(length(unique(pew_core_habitat$habitat)), "Set2")
# hab_colorbrew <- c("#80B1D3", "#FDB462", "#B3DE69", "#FCCDE5", "#8DD3C7", "#BEBADA", "#FFFFB3", "#FB8072") # set 3 

ggplot(habitat_detailed, aes(x=state, y=proportion, fill=habitat_simplified)) +
  geom_bar(stat = "identity") +
  hab_colors +
  facet_wrap(.~sampled_v_area) +
  xlab(NULL) +
  coord_flip() +
  theme_minimal(base_size = 15) +
  # labs(caption = fig_citation, fill = element_blank(), y = "Proportion") +
  ggtitle("State-level Comparison of Sampled and Mapped Habitat") +
  theme(legend.position = "bottom")
  # theme(axis.text.y = element_text(size = 12))

# save figure (with some visual adjustments)
ggsave("blue_carbon_data_inventory/figures/sampled_v_estimated_habitat_2023.jpg", width = 8, height = 6,
       ggplot(habitat_detailed, aes(x=state, y=proportion, fill=habitat_simplified)) +
  geom_bar(stat = "identity") +
  hab_colors +
  facet_wrap(.~sampled_v_area) +
  xlab(NULL) +
  coord_flip() +
  theme_minimal() +
  ggtitle("State-level Comparison of Sampled and Mapped Habitat"))

# state specific breakdown
# compare sampled to mapped habitat representation

# # need to cycle through a for loop to output 
# habitat_detailed %>% 
#   filter(state == "WA") %>%
#   mutate(habitat_simplified = fct_reorder(habitat_simplified, proportion)) %>%
#   # plot
#   ggplot(aes(x = habitat_simplified, y = proportion, fill = habitat_simplified)) +
#   geom_col(position = "dodge", show.legend = F) +
#   facet_grid(~sampled_v_area) +
#   hab_colors +
#   scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
#   coord_flip() +
#   theme_bw() +
#   theme(axis.text.y = element_text(size = 10)) +
#   ylab("Proportion") + xlab("Habitat") +
#   ggtitle("Washington Estimated vs. Sampled Habitat")
```

## Data Inventory Change

To demonstrate the change between the initial and second data inventory, the difference in score was calculated for each of the four metrics (quality, quantity, habitat coverage, and spatial coverage).

```{r state-change, fig.height=12, fig.width=12, eval=FALSE}
# try a state-based visualization instead

# all_metrics_long <- all_metrics %>% 
#   select(state, contains("diff")) %>% 
#   pivot_longer(cols = -state, names_to = "metric", values_to = "change") %>% 
#   mutate(metric = recode(metric, 
#                          "spatial_diff" = "Spatial",
#                          "quant_diff" = "Quantity",
#                          "qual_diff" = "Quality",
#                           "hab_diff" = "Habitat"))
# 
# all_metrics_long %>% 
#   ggplot(aes(metric, change, col = metric)) +
#   geom_point(size = 3) +
#   geom_segment(aes(xend = metric, yend = 0)) +
#   coord_flip() +
#   facet_wrap(~state, dir = "v") +
#   theme_bw(base_size = 20) +
#     theme(legend.position = "bottom")

```

<!-- ### Metric-Based Overview -->

```{r metric-change, fig.height=12, fig.width=12, eval=FALSE}

# all_metrics_long %>% 
#   ggplot(aes(state, change, col = state)) +
#   geom_point(size = 3) +
#   geom_segment(aes(xend = state, yend = 0)) +
#   coord_flip() +
#   facet_wrap(~metric, nrow = 1) +
#   theme_bw(base_size = 20) +
#   theme(legend.position = "none")
```

### Quantity

```{r echo=FALSE, message=FALSE, fig.width=12, fig.height=7}

all_metrics %>% 
  mutate(state = fct_reorder(state, quantity_metric_2022)) %>% 
  select(state, contains("quantity")) %>% 
  pivot_longer(cols = -state, names_to = "metric_iteration", values_to = "score") %>% 
  ggplot() +
  geom_point(aes(score, state, col = metric_iteration), alpha = 0.5, size = 3) +
  geom_segment(data = all_metrics %>% 
                 filter(quant_diff > 0), aes(y = state, yend = state, x = quantity_metric_2022, xend = quantity_metric_2023), 
               colour = "black", arrow = arrow(length = unit(0.2, "cm"))) +
  # facet_wrap(climate_zone ~ management, dir = "h") +
  xlab("Cores per 1000ha Wetland Habitat") + ylab("") +
  theme_bw(base_size = 15) +
  theme(legend.position="bottom") +
  ggtitle("Change in Quantity of Available Cores per State")

ggsave("blue_carbon_data_inventory/figures/change_quantity.jpg", width = 6, height = 6)
```

### Quality

```{r echo=FALSE, message=FALSE, fig.width=12, fig.height=7}

all_metrics %>% 
  mutate(state = fct_reorder(state, quality_metric_2022)) %>% 
  select(state, contains("quality")) %>% 
  pivot_longer(cols = -state, names_to = "metric_iteration", values_to = "score") %>% 
  ggplot() +
  geom_point(aes(score, state, col = metric_iteration), alpha = 0.5, size = 3) +
  geom_segment(data = all_metrics %>% 
                 filter(qual_diff > 0), aes(y = state, yend = state, x = quality_metric_2022, xend = quality_metric_2023), 
               colour = "black", arrow = arrow(length = unit(0.2, "cm"))) +
  # facet_wrap(climate_zone ~ management, dir = "h") +
  xlab("Cores per 1000ha Wetland Habitat") + ylab("") +
  theme_bw(base_size = 15) +
  theme(legend.position="bottom") +
  ggtitle("Change in Use and Completeness of Cores")

ggsave("blue_carbon_data_inventory/figures/change_quality.jpg", width = 6, height = 6)
```

### Spatial Coverage

```{r echo=FALSE, message=FALSE, fig.width=12, fig.height=7}

all_metrics %>% 
  mutate(state = fct_reorder(state, spatial_metric_2022)) %>% 
  select(state, contains("spatial_metric")) %>% 
  pivot_longer(cols = -state, names_to = "metric_iteration", values_to = "score") %>% 
  ggplot() +
  geom_point(aes(score, state, col = metric_iteration), alpha = 0.5, size = 3) +
  geom_segment(data = all_metrics %>% 
                 filter(spatial_diff != 0), aes(y = state, yend = state, x = spatial_metric_2022, xend = spatial_metric_2023), 
               colour = "black", arrow = arrow(length = unit(0.2, "cm"))) +
  # facet_wrap(climate_zone ~ management, dir = "h") +
  xlab("Mean Sampling Distance\n/ Idealized Sampling Distance") + ylab("") +
  theme_bw(base_size = 15) +
  theme(legend.position="bottom") +
  ggtitle("Change in Spatial Coverage of Cores")

ggsave("blue_carbon_data_inventory/figures/change_spatial.jpg", width = 6, height = 6)
```

### Habitat Coverage

```{r echo=FALSE, message=FALSE, fig.width=12, fig.height=7}

all_metrics %>% 
  mutate(state = fct_reorder(state, habitat_metric_2022)) %>% 
  select(state, contains("habitat")) %>% 
  pivot_longer(cols = -state, names_to = "metric_iteration", values_to = "score") %>% 
  ggplot() +
  geom_point(aes(score, state, col = metric_iteration), alpha = 0.5, size = 3) +
  geom_segment(data = all_metrics %>% 
                 filter(hab_diff != 0), aes(y = state, yend = state, x = habitat_metric_2022, xend = habitat_metric_2023), 
               colour = "black", arrow = arrow(length = unit(0.2, "cm"))) +
  # facet_wrap(climate_zone ~ management, dir = "h") +
  xlab("Difference Between Sampled\nand Mapped Proportions") + ylab("") +
  theme_bw(base_size = 15) +
  theme(legend.position="bottom") +
  ggtitle("Change in Habitat Representativeness of Cores")

ggsave("blue_carbon_data_inventory/figures/change_habitat.jpg", width = 6, height = 6)

```

## Database Citation

Coastal Carbon Network (2023). Database: Coastal Carbon Library (Version 1.1.0). Smithsonian Environmental Research Center. Dataset. https://doi.org/10.25573/serc.21565671

